Copulas and their potential for ecology

### 2019-12-14 Paper.Rmd
table added

knitr::kable(tab_data_info, 
             #format="latex", align="l",
             format="latex", align="l",escape=F,
             caption = "Summary table for the data we used. Bold entries are the multivariate data, the rest are bivariate datasets (see section \ref{Data}). Basal metabolic rate=BMR. \\label{tab_data_info}", 
             booktabs = T,linesep = "\\addlinespace") %>%
             column_spec(1:4,width="3.4 cm")%>%
             row_spec(c(4,5,6,8), bold=T)
```

@@

We fit 16 bivariate copula families, exhibiting a variety of 
lower- and upper-tail dependence characteristics, with bivariate datasets. The purpose
of using a large collection of families was solely to include a 
variety of alternative dependence structures to have a robust model
selection and multi-model inference procedure. For that purpose, it 
is not important for the reader to
understand the details of these copulas,
and, additonally, these copulas have been 
described in more detail elsewhere [@Joe1997]; so 
we only briefly identify each family, say a few words about tail dependence 
of the families, and provide pictorial descriptions (Figs 
\ref{figped_N}-\ref{figped_BB1}, \ref{SM-figped_SC} and \ref{SM-figped_G}-\ref{SM-figped_SBB8}). 
The pictorial decriptions can also be used to aid quick comparisons of copula
families when alternative families are being considered for future applications. 
We used several families that can exhibit positive lower-tail dependence 
(of strength depending on parameters) and 
zero upper-tail dependence: the Clayton, survival Gumbel, survival Joe
and survival BB6 families. We used several families that can exhibit positive
upper-tail dependence (strength depends on parameters) and zero lower-tail
dependence: the survival Clayton, Gumbel, Joe and BB6 families. We used
families that show zero upper- and lower-tail dependence: the normal
and Frank families. These families have pdfs symmetric about the line
$v=-u+1$. We also used several families that can show both upper- and
lower-tail dependence, in relative amounts that depend on parameter
values: the BB1, survival BB1, BB7 and survival BB7 families. 
We also used the BB8 family, which shows zero lower-tail dependence,
and zero upper-tail dependence except for a boundary case for the parameters.
And we used the survival BB8 copula, which shows zero upper-tail dependence, 
and zero lower-tail dependence except for a boundary case.
"Survival" families are rotations of the copula with the similar name by
180 degrees. 
See @Joe1997 for details on all these families. We used the implementations
of these families provided in the `VineCopula` package for the R programming
language. See Figs \ref{figped_N}-\ref{figped_BB1}, \ref{SM-figped_SC} and \ref{SM-figped_G}-\ref{SM-figped_SBB8} 
for visual depictions of the pdfs of these copulas and how
pdfs and tail dependence are influenced by the parameters. See 
Table X for a summary of some properties of each family.
<!--***DAN: Shyamolina, please make a table, and replace the X above with 
a citation to that table. Please place the table chunk in such a way that 
this citation becomes table 2, since prior to this point the only main text
table that has been cited is table 1, and we want the table citations to 
come in order. The columns of the table should include: 1) copula family name,
2) number of parameters for the family; 3) info on tail dependence that is
in the above para; 4) a reference to the figure (in the main text or sup mat)
where there is a picture of this family. Anything else? There should be a 
row for each of our 16 copula families.-->

>>>>

knitr::kable(tab_data_info, 
             #format="latex", align="l",
             format="latex", align="l",escape=F,
             caption = "Summary table for the data we used. Bold entries are the multivariate data, the rest are bivariate datasets (see section \\ref{Data}). Basal metabolic rate=BMR. \\label{tab_data_info}", 
             booktabs = T,linesep = "\\addlinespace") %>%
             column_spec(1:4,width="3.4 cm")%>%
             row_spec(c(4,5,6,8), bold=T)
```

```{r tab_summary_copula_info, echo=F, results="asis",message=F}
library(kableExtra)
library(dplyr)
tab_copula_info<-as.data.frame(matrix(NA,nrow=16,ncol=4))
colnames(tab_copula_info)<-c("Copula family","No. of parameters","Stronger dependent tail", "Reference")

tab_copula_info$`Copula family`<-c("Normal",
                                   "Clayton","Survival Clayton",
                                   "Gumbel","Survival Gumbel",
                                   "Joe","Survival Joe",
                                   "Frank", 
                                   "BB1","Survival BB1",
                                   "BB6","Survival BB6", 
                                   "BB7","Survival BB7", 
                                   "BB8","Survival BB8")

tab_copula_info$`No. of parameters`<-c("1",
                                       "1","1",
                                       "1","1",
                                       "1","1",
                                       "1",
                                       "2","2",
                                       "2","2",
                                       "2","2",
                                       "2","2")

tab_copula_info$`Stronger dependent tail`<-c("None",
                                             "Lower","Upper",
                                             "Upper","Lower",
                                             "Upper","Lower",
                                             "None",
                                             "Any (depending on the parameters)","Any (depending on the parameters)",
                                             "Upper","Lower",
                                             "Any (depending on the parameters)","Any (depending on the parameters)",
                                             "Either upper or none (depending on the parameters)","Either lower or none (depending on the parameters)")

# caution: this fig. numbers are not auto-linked
tab_copula_info$Reference<-c("Fig. 4",
                              "Fig. 5","Fig. S3",
                              "Fig. S7","Fig. S8",
                              "Fig. S9","Fig. S10",
                              "Fig. S11",
                              "Fig. 6","Fig. S12",
                              "Fig. S13","Fig. S14",
                              "Fig. S15","Fig. S16",
                              "Fig. S17","Fig. S18")

knitr::kable(tab_copula_info, 
             format="latex", align="l",escape=F,
             caption = "Summary table for the sixteen copula families we used (see section \\ref{Background}). \\label{tab_copula_info}", 
             booktabs = T,linesep = "\\addlinespace") %>%
             column_spec(2,width="2 cm")
```

@@

We fit 16 bivariate copula families, exhibiting a variety of 
lower- and upper-tail dependence characteristics, with bivariate datasets. The purpose
of using a large collection of families was solely to include a 
variety of alternative dependence structures to have a robust model
selection and multi-model inference procedure. For that purpose, it 
is not important for the reader to
understand the details of these copulas,
and, additonally, these copulas have been 
described in more detail elsewhere [@Joe1997]; so 
we only briefly identify each family, say a few words about tail dependence 
of the families, and provide pictorial descriptions (Figs 
\ref{figped_N}-\ref{figped_BB1}, \ref{SM-figped_SC} and \ref{SM-figped_G}-\ref{SM-figped_SBB8}). 
The pictorial decriptions can also be used to aid quick comparisons of copula
families when alternative families are being considered for future applications. 
We used several families that can exhibit positive lower-tail dependence 
(of strength depending on parameters) and 
zero upper-tail dependence: the Clayton, survival Gumbel, survival Joe
and survival BB6 families. We used several families that can exhibit positive
upper-tail dependence (strength depends on parameters) and zero lower-tail
dependence: the survival Clayton, Gumbel, Joe and BB6 families. We used
families that show zero upper- and lower-tail dependence: the normal
and Frank families. These families have pdfs symmetric about the line
$v=-u+1$. We also used several families that can show both upper- and
lower-tail dependence, in relative amounts that depend on parameter
values: the BB1, survival BB1, BB7 and survival BB7 families. 
We also used the BB8 family, which shows zero lower-tail dependence,
and zero upper-tail dependence except for a boundary case for the parameters.
And we used the survival BB8 copula, which shows zero upper-tail dependence, 
and zero lower-tail dependence except for a boundary case.
"Survival" families are rotations of the copula with the similar name by
180 degrees. 
See @Joe1997 for details on all these families. We used the implementations
of these families provided in the `VineCopula` package for the R programming
language. See Figs \ref{figped_N}-\ref{figped_BB1}, \ref{SM-figped_SC} and \ref{SM-figped_G}-\ref{SM-figped_SBB8} 
for visual depictions of the pdfs of these copulas and how
pdfs and tail dependence are influenced by the parameters. See 
Table \ref{tab_copula_info} for a summary of some properties of each family.
<!--***DAN: Shyamolina, please make a table, and replace the X above with 
a citation to that table. Please place the table chunk in such a way that 
this citation becomes table 2, since prior to this point the only main text
table that has been cited is table 1, and we want the table citations to 
come in order. The columns of the table should include: 1) copula family name,
2) number of parameters for the family; 3) info on tail dependence that is
in the above para; 4) a reference to the figure (in the main text or sup mat)
where there is a picture of this family. Anything else? There should be a 
row for each of our 16 copula families.-->
